---
title: "Class 6: R functions"
author: "Kevin (A16482696)"
format: pdf
---

Functions are how we get work done in R. We call functions to do everything from reading data to doing analysis and outputing plots and resutls. 

All functions in R have at least 3 thing:

- a **name** (you get to pick this)
- input **arguments** (there can be only one or loads - again your call)
- the **body**(where the work gets done, this code between the curly brackets)

## A First silly function

Let's write a function to add ome numbers. We can call it `add()`

```{r}
x <- 10
y <- 10
x + y
```

```{r}
add <- function(x){
  y <- 10
  x + y
}
  
```

Can I just use my new function?

```{r}
add(1)
```

Let's make it a bit more flexible

```{r}
add <- function(x,y=1){
  x + y
}

add(x=10,y=10)
```
```{r}
add(10)
```

```{r}
add(10,100)
```

## 2nd example grade() function

Write a function to grade student work. 

We will start with a simple version of the problem and the following example student vectors:

```{r}
# Example input vectors to start with
student1 <- c(100, 100, 100, 100, 100, 100, 100, 90)
student2 <- c(100, NA, 90, 90, 90, 90, 97, 80)
student3 <- c(90, NA, NA, NA, NA, NA, NA, NA)

```

Start with student1
```{r}
min(student1)
```


Google told me about min() and max()

```{r}
min(student1)
```
```{r}
which.min(student1)
```
```{r}
student1[8]
```

```{r}
student1[which.min(student1)]
```

```{r}
student1[-8]
```

Our first working snippet that drops the lowest score and calculates the mean.

```{r}
mean(student1[-which.min(student1)])
```

```{r}
x <- student1
mean(x[-which.min(x)])
```

Our appraoch to the NA problem (missing homeworks): We can replace all NA values with zero. 

1st Taske is find the A values(i.e. where are they in the vector)
```{r}
x <- student2
```

```{r}
x
is.na(x)
```

I have found the NA (TRUE) values from `is.na()` now I want to make them equal to zero (overwrite them/mask them etc.)

```{r}
y <- 1:5
y
y[y>3] <- 0
y
```


I want to combine the `is.na(x)` with making these elements equal to zero. 
And then take this "masked" (vector of student scores with NA values as zero) and drop the lowest and get the mean

```{r}
x[is.na(x)] <- 0
x
```

```{r}
mean(x[-which.min(x)])
```

```{r}
x <- student2
x[is.na(x)] <- 0
mean(x[-which.min(x)])
```

Now I can turn my most awesome snipet into my first function.

```{r}
grade <- function(x){
  # MAKE NA (Missing work) equal to zero
  x[is.na(x)] <- 0
  # Drop lowest scroe and get mean
  mean(x[-which.min(x)])
}
```

```{r}
grade(student3)
```
>Write a function grade() to determine an overall grade from a vector of student homework
assignment scores dropping the lowest single score. If a student misses a homework (i.e. has an
NA value) this can be used as a score to be potentially dropped. Your final function should be
adquately explained with code comments and be able to work on an example class gradebook
such as this one in CSV format: “https://tinyurl.com/gradeinput” [3pts]

```{r}
url <- "https://tinyurl.com/gradeinput"
gradebook <- read.csv(url, row.names = 1)
head(gradebook)
```

The `apply()` function in R is super useful but can be little confusing to begin with. Lets have a look how it works

```{r}
ans <- apply(gradebook, 1, grade)
ans
```


> Using your grade() function and the supplied gradebook, Who is the top scoring student
overall in the gradebook? [3pts]

```{r}
which.max(ans)
max(ans)
```

>From your analysis of the gradebook, which homework was toughest on students (i.e. obtained
the lowest scores overall?

```{r}
which.min(apply(gradebook, 2, mean, na.rm=TRUE))
```

> Optional Extension: From your analysis of the gradebook, which homework was most
predictive of overall score (i.e. highest correlation with average grade score)? [1pt]

```{r}
#ans
cor(gradebook$hw1,ans)

```

```{r}
cor(gradebook$hw5,ans)
```

Mask all the NA values to zero in gradebook
```{r}
mask <- gradebook
mask[is.na(mask)] <- 0
mask
```

```{r}
cor(mask$hw2, ans)
```

Now we can use `apply()` to examine the correlatino of every assignment in the masked gradebook to the overall score for each student in the class

```{r}
apply(mask, 2, cor, y=ans)
```

> Make sure you save your Quarto document and can click the “Render” (or Rmarkdown”Knit”) button to generate a PDF foramt report without errors. Finally, submit your PDF
to gradescope. [1pt]































